---
title: "Preprocessing"
author: "Victor G."
date: "2 de diciembre de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
setwd("E:/Documents/Universidad (FIB)/APA/practica")
superConductor <- read.csv("train.csv", na.strings = c("?"))
superConductorCompo <- read.csv("unique_m.csv", na.strings = c("?"))
```

Generating full dataframe
```{r}
df <- data.frame(superConductor, superConductorCompo)#merge(superConductor, superConductorCompo)
drops <- c("critical_temp.1", "material")
df<-df[,!(names(df) %in% drops)]
t(head(df))
```


```{r}
#options(max.print=1000000)
t(summary(df))
```

From this we see Neon, Helium, Argon, Kripton, Xenon, Polonium, Astatine and Radon are constant at 0. We will drop them.

```{r}
#drop neon, helium, Argon, kripton, Xenon, , Polonium, Astatine and Radon
drops <- c("Ne", "He", "Ar", "Kr", "Xe","Po", "At", "Rn")
df<-df[,!(names(df) %in% drops)]
```

We consider dropping Cadmium too, since it looks like there might not be too many cases, but further inspection tells us this is not the case, as it has 178 non-0 rows.

We know there are no NAs from the way data was generated. We need only do feature extraction and normalisation / log transforms.

```{r}
#Very time consuming code, only shows slightly skewed graphics for some of them
for (col in names(superConductor)){
  hist(df[,col], main = paste("Histogram of ", col))
}

# Some seem to need a log
```

We will center on correcting leptokurtic and positively skewed distributions, and we will do so applying logarithmic transformations.

```{r}
library(moments)
for (col in names(superConductor)){ #only to physical variables, not to proportions. Too many 0s in them, distorted.
  if(skewness(df[,col]) > 2 && kurtosis(df[,col]) > 2) { # We use 2 as threshold for logs
      if (min(df[,col])==0) {
        print(paste0(col, " needs log1"))
      } else {
        print(paste0(col, " needs log"))
      }
  }
}

```

```{r}
log <- c("mean_Density","wtd_mean_Density","mean_FusionHeat","wtd_mean_FusionHeat","gmean_FusionHeat","wtd_gmean_FusionHeat","gmean_ThermalConductivity","wtd_gmean_ThermalConductivity")
log1 <- c("wtd_range_atomic_mass","wtd_range_Density","range_FusionHeat","wtd_range_FusionHeat","std_FusionHeat","wtd_std_FusionHeat")

for (col in log) {
  df[,paste0(col, "_log")] = log(df[,col])
}

for (col in log1) {
  df[,paste0(col, "_log1")] = log1p(df[,col])
}
```

```{r}
t(summary(df))
```

Now we have some variables with a more correct kurtosis. Let's check one of them:

```{r}
hist(df[,"mean_Density"], main = "Histogram of mean_Density")
hist(df[,"mean_Density_log"], main = "Histogram of mean_Density_log")
```

As we can see, much more normal and symmetric. We drop the old ones

```{r}
df <- df[, -which(names(df) %in% c(log,log1))]
```

Now we should standardize, it is needed for methods such as KNN, and it doesn't affect negatively in general

```{r}
standardized <- as.data.frame(scale(df))
```

Finally, sampling and splitting the data.

```{r}
set.seed(42)
learn_size <- floor(0.7*nrow(standardized))
learn_ind <- sample(seq_len(nrow(standardized)), size = learn_size)
learn <- standardized[learn_ind,]
test <- standardized[-learn_ind,]
```

And save

```{r}
write.csv(learn, file="learn_set.csv")
write.csv(test, file="test_set.csv")

```